---
title: "Homework 3"
author: "Neal Marquez"
date: "October 23, 2017"
output: pdf_document
---

## Libraries

```{r warning=FALSE,message=FALSE,error=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(pander)
library(ggplot2)
library(stringr)
```

## Utilities

```{r}
acs_url <- "https://api.census.gov/data/2015/acs5"
wb_url <- "http://api.worldbank.org"
load("./tokens.Rdata")
fb_url <- "https://graph.facebook.com/v2.10"
fb_ads_url <- str_c(fb_url, paste0("act_", buisiness_id), 
                    "reachestimate", sep = "/")

process_wb_data <- function(d) {
  as_data_frame(d) %>%
    unnest() %>%
    select(-decimal) %>%
    group_by(value, date) %>%
    summarise_all(first) %>%
    ungroup()
}

admin1_df <- function(iso2){
  
  wb_query_pg1 <- list(page=1, format="json")
  r_pop_pg1 <- 
    GET(wb_url, query = wb_query_pg1,
        path=paste0("countries/", iso2, "/indicators/SP.POP.TOTL"))
  
  wb_query_pg2 <- list(page=2, format="json")
  r_pop_pg2 <- 
    GET(wb_url, query=wb_query_pg2,
        path=paste0("countries/", iso2, "/indicators/SP.POP.TOTL"))

  df_pg1 <-
    map(content(r_pop_pg1, as = "parsed")[[2]], 
        process_wb_data) %>% 
    bind_rows()

  df_pg2 <- 
    map(content(r_pop_pg2, as = "parsed")[[2]], 
        process_wb_data) %>% 
    bind_rows()

  r_pop_df <- 
    bind_rows(df_pg2, df_pg1) %>%
    mutate(value = as.numeric(value), 
           date = as.numeric(date))

  r_pop_df
}

country_result <- function(iso2){
  targeting_specs <- paste0('{"geo_locations": {"countries": ["', iso2, '"]}}')
  
  fb_query_multiple <- list(
    access_token = fb_token, 
    currency = "USD", 
    optimize_for = "NONE", 
    targeting_spec = targeting_specs
  )
  
  r <- GET(fb_ads_url, query = fb_query_multiple)
  
  return(data.frame(Country=iso2, Population=fromJSON(prettify(r))$data$users))
}
```

1. Use the 2016 1-year ACS to get the population of every state. What are the 5 smallest states?

```{R}
acs1_url <- "https://api.census.gov/data/2016/acs/acs1" 

acs1_query <- list(
  get = "B01003_001E,B01003_001M,NAME",
  `for` = "state:*",
  key = census_token
)

acs1_r <- GET(url = acs1_url, query = acs1_query)

acs1_m <- fromJSON(content(acs1_r, as = "text"))
acs1_df <- as_data_frame(acs1_m[-1, ])
colnames(acs1_df) <- acs1_m[1, ]

acs1_df <- 
  acs1_df %>%  
  mutate(B01003_001E = as.numeric(B01003_001E), 
         B01003_001M = as.numeric(B01003_001M)) %>%
  arrange(B01003_001E) %>%
  select(-B01003_001M) %>%
  rename(N = B01003_001E) %>% filter(state != 11 & state != 72) # no DC or PR

pander(head(acs1_df, n=5))
```

2. Request population indicators for three other countries from the World Bank API and plot their population change over time.

```{r}
iso2s <- c("mx", "gt", "sv", "ni", "cr")
multi_country_df <- do.call(rbind, lapply(iso2s, function(x) admin1_df(x)))

ggplot(multi_country_df,
       aes(x=date, y=log(value), group=country, color=country)) + 
  geom_line() + 
  theme_minimal(base_size=20) + 
  labs(title="Log Population of Latin America, 1960-2016", x=NULL, y=NULL) + 
  scale_color_discrete("Country")
```

4. Pick 3 other countries and report the number of Facebook users in each. Compare these numbers to the actual populations from the World Bank or some other source.

```{r}
iso2s <- c("US", "CA", "MX")
facepop <- do.call(rbind, lapply(iso2s, country_result)) %>%
  mutate(Country=as.character(Country))
multi_country_df <- do.call(rbind, lapply(tolower(iso2s), function(x) 
  admin1_df(x))) %>% filter(date == 2016)

multi_country_df %>% rename(Country=country) %>% 
  left_join(facepop, "Country") %>%
  ggplot(aes(x=value, y=Population, color=Country)) + geom_point() + 
  xlab("World Bank Population Est.") + ylab("Facebook User Population") +
  labs(title="")
```

5. Pick another US state and age range, and compare the numbers of men and women.

```{r}
ts1 <- "{\n\"geo_locations\":{\n\"regions\":[{\n\"key\":\"3893\",\n\"name\":\"Wyoming\"\n}]\n},\n\"genders\":[1],\n\"age_min\": 15,\n\"age_max\":25\n}\n"
ts2 <- "{\n\"geo_locations\":{\n\"regions\":[{\n\"key\":\"3893\",\n\"name\":\"Wyoming\"\n}]\n},\n\"genders\":[2],\n\"age_min\": 15,\n\"age_max\":25\n}\n"

process_fb_response <- function(ts) {
  Sys.sleep(5)
  
  fb_query <- list(
    access_token = fb_token, 
    currency = "USD", 
    optimize_for = "NONE", 
    targeting_spec = minify(ts)
  )
  
  r <- GET(fb_ads_url, query = fb_query)
  
  ts_df <- 
    fromJSON(ts) %>%
    as_data_frame() %>%
    unnest(geo_locations)
  
  r_df <- 
    content(r, as = "parsed")$data %>%
    as_data_frame() %>%
    select(users)
  
  bind_cols(ts_df, r_df)
}

dfWY <- rbind(process_fb_response(ts1), process_fb_response(ts2))
dfWY
```

6. Get the Facebook user population for each US state. Compare these estimates to population estimates from the Census ACS.

```{r}
usvec <- readLines(paste0(
  "./r-estimates-fb-ads-master/targeting_specs/",
  "targeting_spec_us_states.json"))

fbstatedf <- do.call(rbind, lapply(usvec, process_fb_response))

fbstatedf %>% rename(NAME=name) %>% right_join(acs1_df, "NAME") %>%
  ggplot(aes(x=N, y=users)) + geom_point() +
  labs(x="Population(ACS)", y="FB Users", title="State Population vs FB Users")
```